%
%   This program is free software: you can redistribute it and/or modify
%   it under the terms of the GNU General Public License as published by
%   the Free Software Foundation, either version 3 of the License, or
%   (at your option) any later version.
%
%   This program is distributed in the hope that it will be useful,
%   but WITHOUT ANY WARRANTY; without even the implied warranty of
%   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%   GNU General Public License for more details.
%
%   You should have received a copy of the GNU General Public License
%   along with this program.  If not, see <http://www.gnu.org/licenses/>.
%

% Version: $Revision$

For most users, the existing WEKA framework will be sufficient to perform the
task at hand, offering a wide range of filters, classifiers, clusterers, etc.
Researchers, on the other hand, might want to add new algorithms and compare
them against existing ones. The framework with its existing algorithms is not
set in stone, but basically one big plugin framework. With WEKA's automatic
discovery of classes on the classpath, adding new classifiers, filters, etc. to
the existing framework is very easy.

Though algorithms like clusterers, associators, data generators and
attribute selection are not covered in this chapter, their implemention is very
similar to the one of implementing a classifier. You basically choose a
superclass to derive your new algorithm from and then implement additional
interfaces, if necessary. Just check out the other algorithms that are already
implemented.

The section covering the GenericObjectEditor (see chapter
\ref{genericobjecteditor}) shows you how to tell WEKA where to find your
class(es) and therefore making it/them available in the GUI
(Explorer/Experimenter) via the GenericObjectEditor.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Writing a new Classifier %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\section{Writing a new Classifier}
\input{extending_classifier}

%%%%%%%%%%%%%%%%%%%%%%%%
% Writing a new Filter %
%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Writing a new Filter}
\input{extending_filter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Writing other algorithms %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Writing other algorithms}
\input{extending_other_algorithms}

%%%%%%%%%%%%%%%%%%%%%%%%%%
% Extending the Explorer %
%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Extending the Explorer}
\input{extending_explorer}

\newpage
\section{Extending the Knowledge Flow}
\input{extending_knowledgeflow}
